---
title: "fireSense LCC flammability"
author: "Alex Chubaty and Ian Eddy"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

<!-- based on similar vignette in 'fireSenseUtils' package -->

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  eval = FALSE
)
```

## Packages

```{r init}
library(data.table)
library(fasterize)
library(ggplot2)
library(raster)
library(reproducible)
library(sf)

dPath <- file.path("inputs")
```

```{r fire_polys}
fireUrl <- "https://cwfis.cfs.nrcan.gc.ca/downloads/nfdb/fire_poly/current_version/NFDB_poly.zip"
firePolygons <- prepInputs(url = fireUrl, destinationPath = dPath, fun = "sf::read_sf")
```

## Compare LCC and fire data

### LCC2005

```{r LCC2005}
lcc2005Url <- paste0("ftp://ftp.ccrs.nrcan.gc.ca/ad/NLCCLandCover/",
                     "LandcoverCanada2005_250m/LandCoverOfCanada2005_V1_4.zip")
lcc2005 <- prepInputs(url = lcc2005Url, destinationPath = dPath,
                      targetFile = "LCC2005_V1_4a.tif", alsoExtract = NA)
firePolygonsLCC2005 <- sf::st_transform(firePolygons, crs = crs(lcc2005))

## rasterize firePolygons
firePolygonsLCC2005$dummyVar <- 1
firePolygonsLCC2005 <- firePolygonsLCC2005[firePolygonsLCC2005$YEAR > 2004, ]
fireRasterLCC2005 <- fasterize(sf = firePolygonsLCC2005, raster = lcc2005, field = "dummyVar")
fireLoc2005 <- 1:ncell(lcc2005)
fireLoc2005 <- fireLoc2005[!is.na(getValues(fireRasterLCC2005))]

burnedLCC2005 <- data.table(pixelID = 1:ncell(lcc2005), lcc = getValues(lcc2005))
burnedLCC2005[pixelID %in% fireLoc2005, burn := 1]

burnCalcLCC2005 <- burnedLCC2005[, .(available = .N, burned = sum(burn, na.rm = TRUE)), .(lcc)]
burnCalcLCC2005[, "percentBurned" := round(burned/available * 100, digits = 3)]
setkey(burnCalcLCC2005, lcc)
burnCalcLCC2005
```

```{r LCC2005_plot, echo=FALSE}
g1 <- ggplot(data = burnCalcLCC2005, aes(x = lcc, y = percentBurned)) + 
  geom_bar(stat = "identity") + 
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  ylab("percent burned (%)") + 
  scale_x_continuous("LCC", labels = as.character(burnCalcLCC2005$lcc), breaks = burnCalcLCC2005$lcc) + 
  geom_hline(yintercept = 1, col = "darkgrey") +
  theme_bw()

gfile1 <- "outputs/AOU/figures/LCC_flammability.png"
ggsave(g1, filename = gfile1)
googledrive::drive_put(gfile1, googledrive::as_id("1DWOgy-XxZO9pmgfRXEzHJPX7jU4x3Vki"))

g1
```

#### LCC2005 Results

```{r LCC2005_nonFlamm, eval = FALSE}
## original fireSense_dataPrepFit defaults
LCC2005_nonFlam <- c(0, 25, 30, 33, 36, 37, 38, 39)
```

### Far North LCC

```{r LCC_FN}
## FAR NORTH LANDCOVER (620 MB)
## unable to download directly b/c of SSL, time outs, and other server problems
##   https://ws.gisetl.lrc.gov.on.ca/fmedatadownload/Packages/FarNorthLandCover.zip
##
source("modules/Ontario_preamble/R/prepInputsFarNorthLCC.R")
farNorthLCCfn <- file.path("inputs", "FarNorth_LCC_Class_RoFL.tif")
if (!file.exists(farNorthLCCfn)) {
  #this file is incorrect - we only want the RoF portion, 
  farNorthLCC <- prepInputsFarNorthLCC(dPath = dPath, filename2 = farNorthLCCfn)
  # farNorthLCC <- postProcess(farNorthLCC, studyArea = <the RoF study area large)
} else {
  farNorthLCC <- raster(farNorthLCCfn)
}


```

### Ecozones

```{r ecozones}
ecozones <- prepInputs(
    url = "https://sis.agr.gc.ca/cansis/nsdb/ecostrat/zone/ecozone_shp.zip",
    targetFile = "ecozones.shp",
    alsoExtract = "similar",
    fun = "sf::st_read",
    destinationPath = dPath,
    targetCRS = crs(farNorthLCC)
  )
ecozones[["ZONE_NAME"]] <- toupper(ecozones[["ZONE_NAME"]])
ecozones <- fasterize(ecozones, raster = farNorthLCC, fun = "last", field = "ZONE_ID")


#TODO: source the NFI 1985-2015 ROF file and finish the analysis


```


```{get Wildfire 1985-2015}
#this is a locally hosted private dataset - the NFI wildfire year raster representing 1985-2015 
#It is ONLY the extent of the Ring of Fire area
wildfire2020 <- prepInputs(
  url = "https://drive.google.com/file/d/1Vc4cOY1jOS1y8P20S14nYBJWwkRj_SPL/",
  targetFile = "Fire_1985-2020_ROF.dat",
  fun = "raster::raster",
  rasterToMatch = farNorthLCC,
  alsoExtract = c("Fire_1985-2020_ROF.dat.ovr",
                  "Fire_1985-2020_ROF.dat.aux.xml",
                  "Fire_1985-2020_ROF.dat.vat.cpg",
                  "Fire_1985-2020_ROF.hdr"),
  destinationPath = dPath
)

```


```{build fire summary table}

farNorthClasses <- c("Clear Open Water" = 1, "Turbid Water" = 2, "Intertidal Mudflat" = 3,
                     "Intertidal Marsh" = 4, "Supratidal Marsh" = 5, "Fresh Water Marsh" = 6,
                     "Heath" = 7, "Thicket Swamp" = 8, "Coniferous Swamp" = 9, "Deciduous Swamp" = 10,
                     "Open Fen" = 11, "Treed Fen" = 12, "Open Bog" = 13, "Treed Bog" = 14,
                     "Sparse Treed" = 15, "Deciduous Treed" = 16, "Mixed Treed" = 17, "Coniferous Treed" = 18,
                     "Disturbance - Non and sparse-woody" = 19, "Disturbance - Treed or shrub" = 20,
                     "Sand/Gravel/Mine Tailings" = 21, "Bedrock" = 22, "Community/Infrastructure" = 23,
                     "Agriculture" = 24, "Cloud/Shadow" = -9, "Other" = -99)
lcc <- data.table("lcc" = farNorthClasses, "lccName" = names(farNorthClasses))

#ecozone 20 is the plains, ecozone 19 is the shield
fireSummaryDT <- data.table(fireYear = getValues(wildfire2020),
                            lcc = getValues(farNorthLCC),
                            ecozone = getValues(ecozones))

post2015 <- fireSummaryDT[lcc, on = "lcc"]
post2015 <- post2015[!lcc %in% c(1,2,3,4,5,6,7,21,22,23,24,-9, -99)]
#calculate the number of unburned and burned pixels - ignoring those that already burned
post2015 <- post2015[ecozone %in% 19:20,]

post2015[, Available := .N, .(lcc, ecozone)]
post2015 <- post2015[fireYear > 2015 & !is.na(fireYear)]
post2015 <- post2015[, .(Burned = .N), .(lcc, lccName, Available, ecozone)]
post2015[, percentBurn := round(Burned/Available * 100, digits = 4)]

post2015Plain <- post2015[ecozone == 20] 
post2015Shield <- post2015[ecozone == 19]
AllBurns <- post2015[, .(Available = sum(Available), Burned = sum(Burned)), .(lccName)]
AllBurns[, percentBurn := round(Burned/Available * 100, digits =4 )]

#we chose to group Bog and Coniferous swamp as BogSwamp (burn > 0.1%  in 5 years)
#FenPlus is all categories where < 0.1% burned over 5 years - 
#Trees are separate. Disturbance gets resampled but defaults to Coniferous Treed based on visual evidence



```




```{r LCCFN_plot, echo=FALSE}
g2 <- ggplot(data = burnCalcFN, aes(x = lcc, y = percentBurned)) + 
  geom_bar(stat = "identity") + 
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  ylab("percent burned (%)") + 
  scale_x_continuous("LCC", labels = as.character(burnCalcFN$lcc), breaks = burnCalcFN$lcc) + 
  geom_hline(yintercept = 1, col = "darkgrey") +
  theme_bw()

gfile2 <- "outputs/ROF/figures/LCC_flammability.png"
ggsave(g2, filename = gfile2)
googledrive::drive_put(gfile2, googledrive::as_id("1OjTkQVUhVq65YPGGOpijZ1ifeRWCwBA4"))

g2
```

#### Far North LCC Results

| LCC | Description                      | LandR        | fireSense     |
|-----|----------------------------------|--------------|---------------|
| 1   | clear open water                 | non-forest   | non-flammable |
| 2   | turbid water                     | non-forest   | non-flammable |
| 3   | intertidal mudflat               | non-forest   | non-flammable |
| 4   | intertidal marsh                 | non-forest   | non-flammable |
| 5   | supertidal marsh                 | non-forest   | non-flammable |
| 6   | freshwater marsh                 | non-forest   | non-flammable |
| 7   | heath                            | non-forest   | fen+          |
| 8   | thicket swamp                    | non-forest   | fen+          |
| 9   | conifer swamp                    | forest       | bogswamp      |
| 10  | deciduous swamp                  | forest       | fen+          |
| 11  | open fen                         | non-forest   | fen+          |
| 12  | treed fen                        | forest       | fen+          |
| 13  | open bog                         | non-forest   | bogswamp      |
| 14  | treed bog                        | forest       | bogswamp      |
| 15  | sparse treed                     | forest       | forest        |
| 16  | deciduous treed                  | forest       | forest        |
| 17  | mixed treed                      | forest       | forest        |
| 18  | conifer treed                    | forest       | forest        |
| 19  | disturbance non and sparse woody | reclassify\* | reclassify\*  |
| 20  | disturbance tree/shrub           | reclassify\* | reclassify\*  |
| 21  | sand/gravel/mine tailings        | non-forest   | non-flammable |
| 22  | bedrock                          | non-forest   | non-flammable |
| 23  | community/infrastructure         | non-forest   | non-flammable |
| 24  | agriculture                      | non-forest   | non-flammable |
