---
title: "fireSense LCC flammability"
author: "Alex Chubaty and Ian Eddy"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

<!-- based on similar vignette in 'fireSenseUtils' package -->

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  eval = FALSE
)
```

## Packages

```{r init}
library(data.table)
library(fasterize)
library(ggplot2)
library(raster)
library(reproducible)
library(sf)

dPath <- file.path("inputs")
```

```{r fire_polys}
fireUrl <- "https://cwfis.cfs.nrcan.gc.ca/downloads/nfdb/fire_poly/current_version/NFDB_poly.zip"
firePolygons <- prepInputs(url = fireUrl, destinationPath = dPath, fun = "sf::read_sf")
```

## Compare LCC and fire data

### LCC2005

```{r LCC2005}
lcc2005Url <- paste0("ftp://ftp.ccrs.nrcan.gc.ca/ad/NLCCLandCover/",
                     "LandcoverCanada2005_250m/LandCoverOfCanada2005_V1_4.zip")
lcc2005 <- prepInputs(url = lcc2005Url, destinationPath = dPath,
                      targetFile = "LCC2005_V1_4a.tif", alsoExtract = NA)
firePolygonsLCC2005 <- sf::st_transform(firePolygons, crs = crs(lcc2005))

## rasterize firePolygons
firePolygonsLCC2005$dummyVar <- 1
firePolygonsLCC2005 <- firePolygonsLCC2005[firePolygonsLCC2005$YEAR > 2004, ]
fireRasterLCC2005 <- fasterize(sf = firePolygonsLCC2005, raster = lcc2005, field = "dummyVar")
fireLoc2005 <- 1:ncell(lcc2005)
fireLoc2005 <- fireLoc2005[!is.na(getValues(fireRasterLCC2005))]

burnedLCC2005 <- data.table(pixelID = 1:ncell(lcc2005), lcc = getValues(lcc2005))
burnedLCC2005[pixelID %in% fireLoc2005, burn := 1]

burnCalcLCC2005 <- burnedLCC2005[, .(available = .N, burned = sum(burn, na.rm = TRUE)), .(lcc)]
burnCalcLCC2005[, "percentBurned" := round(burned/available * 100, digits = 3)]
setkey(burnCalcLCC2005, lcc)
burnCalcLCC2005
```

```{r LCC2005_plot, echo=FALSE}
g1 <- ggplot(data = burnCalcLCC2005, aes(x = lcc, y = percentBurned)) + 
  geom_bar(stat = "identity") + 
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  ylab("percent burned (%)") + 
  scale_x_continuous("LCC", labels = as.character(burnCalcLCC2005$lcc), breaks = burnCalcLCC2005$lcc) + 
  geom_hline(yintercept = 1, col = "darkgrey") +
  theme_bw()

gfile1 <- "outputs/AOU/figures/LCC_flammability.png"
ggsave(g1, filename = gfile1)
googledrive::drive_put(gfile1, googledrive::as_id("1DWOgy-XxZO9pmgfRXEzHJPX7jU4x3Vki"))

g1
```

#### LCC2005 Results

```{r LCC2005_nonFlamm, eval = FALSE}
## original fireSense_dataPrepFit defaults
LCC2005_nonFlam <- c(0, 25, 30, 33, 36, 37, 38, 39)
```

### Far North LCC

```{r LCC_FN}
## FAR NORTH LANDCOVER (620 MB)
## unable to download directly b/c of SSL, time outs, and other server problems
##   https://ws.gisetl.lrc.gov.on.ca/fmedatadownload/Packages/FarNorthLandCover.zip
##

FarNorthLandCoverZip <- file.path(dPath, "FarNorthLandCover.zip")
rFiles <- c("New folder/Class/FarNorth_LandCover_Class_UTM15.tif",
            "New folder/Class/FarNorth_LandCover_Class_UTM16.tif",
            "New folder/Class/FarNorth_LandCover_Class_UTM17.tif")

if (!file.exists(FarNorthLandCoverZip)) {
  googledrive::drive_download(file = as_id("1_QeOAeSHjOjEU1iTqACRj8BYo3ysQKUi"),
                              path = FarNorthLandCoverZip)
  archive::archive_extract(
    archive = file.path(dPath, "FarNorthLandCover.zip"),
    dir = file.path(dPath, "FarNorthLandCover"),
    files = "Version 1.4/TIF Format/ab1da0f2-7bba-430b-af11-d503865ff130-TIF.zip"
  )

  archive::archive_extract(
    archive = file.path(dPath, "FarNorthLandCover", "Version 1.4", "TIF Format",
                        "ab1da0f2-7bba-430b-af11-d503865ff130-TIF.zip"),
    dir = file.path(dPath, "FarNorthLandCover"),
    files = rFiles
  )

  vapply(file.path(dPath, "FarNorthLandCover", rFiles), file.copy, logical(1),
         to = file.path(dPath, "FarNorthLandCover"))
  unlink(file.path(dPath, "FarNorthLandCover", "New folder"), recursive = TRUE)
  unlink(file.path(dPath, "FarNorthLandCover", "Version 1.4"), recursive = TRUE)
}

r15 <- raster(file.path(dPath, "FarNorthLandCover", basename(rFiles))[1])
r16 <- raster(file.path(dPath, "FarNorthLandCover", basename(rFiles))[2])
r17 <- raster(file.path(dPath, "FarNorthLandCover", basename(rFiles))[3])

burnCalcFN <- lapply(list(r15, r16, r17), function(r) {
  r[r[] > 24] <- NA_integer_ ## a few pixels have values >24, so make them NA
  
  firePolygons2 <- sf::st_transform(firePolygons, crs = crs(r))
  
  ## rasterize firePolygons
  firePolygons2$dummyVar <- 1
  firePolygons2 <- firePolygons2[firePolygons2$YEAR > 2004, ]
  fireRaster <- fasterize(sf = firePolygons2, raster = r, field = "dummyVar")
  fireLoc <- 1:ncell(r)
  fireLoc <- fireLoc[!is.na(getValues(fireRaster))]
  
  burned <- data.table(pixelID = 1:ncell(r), lcc = getValues(r))
  burned[pixelID %in% fireLoc, burn := 1]
  
  burnCalc <- burned[, .(available = .N, burned = sum(burn, na.rm = TRUE)), .(lcc)]
  setkey(burnCalc, lcc)
  burnCalc <- na.omit(burnCalc)
  burnCalc
}) %>%
  rbindlist()

burnCalcFN <- burnCalcFN[, lapply(.SD, sum), by = lcc]
setkey(burnCalc, lcc)

burnCalcFN[, "percentBurned" := round(burned/available * 100, digits = 3)]
burnCalcFN
```

```{r LCCFN_plot, echo=FALSE}
g2 <- ggplot(data = burnCalcFN, aes(x = lcc, y = percentBurned)) + 
  geom_bar(stat = "identity") + 
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  ylab("percent burned (%)") + 
  scale_x_continuous("LCC", labels = as.character(burnCalcFN$lcc), breaks = burnCalcFN$lcc) + 
  geom_hline(yintercept = 1, col = "darkgrey") +
  theme_bw()

gfile2 <- "outputs/ROF/figures/LCC_flammability.png"
ggsave(g2, filename = gfile2)
googledrive::drive_put(gfile2, googledrive::as_id("1OjTkQVUhVq65YPGGOpijZ1ifeRWCwBA4"))

g2
```

#### Far North LCC Results

| LCC | Description                      | Flammable? |
|-----|----------------------------------|------------|
| 1   | clear open water                 | N          |
| 2   | turbid water                     | N          |
| 3   | intertidal mudflat               | N          |
| 4   | intertidal marsh                 | N          |
| 5   | supertidal marsh                 | N          |
| 6   | freshwater marsh                 | N          |
| 7   | heath                            | N          |
| 8   | thicket swamp                    | Y          |
| 9   | conifer swamp                    | Y          |
| 10  | deciduous swamp                  | N          |
| 11  | open fen                         | N          |
| 12  | treed fen                        | Y\*        |
| 13  | open bog                         | Y\*        |
| 14  | treed bog                        | Y\*        |
| 15  | sparse treed                     | Y          |
| 16  | deciduous treed                  | Y          |
| 17  | mixed treed                      | Y          |
| 18  | conifer treed                    | Y          |
| 19  | disturbance non and sparse woody | Y          |
| 20  | disturbance tree/shrub           | Y          |
| 21  | sand/gravel/mine tailings        | N\*        |
| 22  | bedrock                          | N\*        |
| 23  | community/infrastructure         | N          |
| 24  | agriculture                      | N          |
